---
title: "Assignment 1: Task 1"
author: "Sophia Leiker"
date: "1/16/2022"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(patchwork)
library(cowplot)
```

## 1. Introduction

This report will explore amphibian abundance data recorded by the Sierra Lakes Inventory Project. From the Environmental Data Initiative repository: “The Sierra Lakes Inventory Project (SLIP) was a research endeavor that ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. We described the physical characteristics of and surveyed aquatic communities for >8,000 lentic water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows.”

The data used comes from: [Sierra Amphibians Dataset](https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3) and metadata, including methods used for data collection can be found here [The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California](https://portal.edirepository.org/nis/metadataviewer?packageid=edi.577.2).


```{r}
# Reading in the data
sierra <- read_csv(here("data", "sierra_amphibians.csv")) %>% 
  clean_names()

```

***

# Part 1

## A. Data Wrangling 
- Selecting Mountain yellow-legged frog (Rana muscosa) --> RAMU
- Finding count each year across all water bodies by life stage
- Exclude 'EggMass' level (find the total number of adult, subadult, and tadpole yellow-legged frogs observed in the entire study region by life stage and year)

```{r filtering}
# filtering to only include certain species of interest
#using lubricate to change survery_date to an actual date 
#only selecting adult, subadult and tadpole

sierra_yellow <- sierra %>% 
  filter(amphibian_species %in% c("RAMU")) %>% 
  mutate(survey_date = mdy(survey_date)) %>% 
  mutate(survey_date = year(survey_date)) %>% 
  filter(amphibian_life_stage %in% c("SubAdult", "Tadpole", "Adult"))
```

## B. Data Analysis
- Grouping by life stage and survey date

```{r}
sierra_count <- sierra_yellow %>% 
  group_by(amphibian_life_stage, survey_date) %>% 
  summarize(total_count = n())

#head(sierra_count)

```

## C. Data visualization

```{r}
graph1 <- ggplot(data = sierra_count, aes(x=survey_date, y= total_count, fill= amphibian_life_stage)) +
  geom_col() +
  labs(x = "Year", y = "Total Count", title = "Amphibian Life Stage for Rana Muscosa in Sierra Lakes Inventory Project", subtitle = "Data from: Knapp, R.A. 2020. The Sierra Lakes Inventory Project") +
  scale_color_manual(values = c("skyblue1","slateblue1","royalblue4")) +
  facet_wrap(~ amphibian_life_stage) +
  theme_minimal() +
  scale_fill_discrete(name = "Life Stage") +
  scale_fill_manual(values = c("dodgerblue1", "mediumaquamarine", "darkorchid3"))

graph1
```

**Figure 1**. Amphibian life stage of Rana Muscosa from 1995 to 2002 broken down by life stage

***

# Part 2

## A. Data Analysis
- Total counts of combined adult and sub-adult frogs in the 5 lakes with greatest total counts
- 5 columns with lake ID

```{r}
sierra_yellow2 <- sierra_yellow %>% 
  filter(amphibian_life_stage %in% c("SubAdult", "Adult")) %>% 
  group_by(lake_id) %>% 
  summarize(total_count = n())

sierra_top5 <- sierra_yellow2 %>% 
  slice_max(total_count, n = 5) %>% #the 6th value is lower than the 5th, so there are no ties between the 5th/6th
  arrange(desc(total_count)) %>% 
  mutate(lake_id = as.character(lake_id))

# Adding "Lake" to the beginning of all the lake names
sierra_top5[1, "lake_id"] <- "Lake 50183"
sierra_top5[2, "lake_id"] <- "Lake 70583"
sierra_top5[3, "lake_id"] <- "Lake 10226"
sierra_top5[4, "lake_id"] <- "Lake 41322"
sierra_top5[5, "lake_id"] <- "Lake 50219"



```

## B. Data Visualization
```{r}
graph2 <- ggplot(data = sierra_top5, aes(x= fct_reorder(lake_id, total_count), y = total_count)) + 
  geom_col(aes(fill = lake_id)) + 
  labs(x = 'Lake ID', y = 'Total Count') +
  #coord_flip() + #flips so columns go from vertical alignment to horizontal alginment
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "Lake ID", y = "Total Count", title = "Lakes with Top 5 Combined Adult and Subadult Rana Muscosa Counts", subtitle = "Data from: Knapp, R.A. 2020. The Sierra Lakes Inventory Project") +
  scale_fill_manual(values = c("slategrey", "steelblue4", "slateblue", "steelblue3", "slateblue4"))

graph2
```

**Figure 2**. Lakes from The Sierra Lakes Inventory Project the with Top 5 Total Observed Counts of Combined Adult and Subadult Rana Muscosa

***

# Part 3 : Combine Graphs!
- Using `cowplot` to combine graph 1 and graph 2

## Assessment of Rana Muscosa Across by Life Stage, Year and Lake

```{r}
#editing graphs to change the placement of things and remove subtitles 
graph3 <- graph1 <- ggplot(data = sierra_count, aes(x=survey_date, y= total_count, fill= amphibian_life_stage)) +
  geom_col() +
  labs(x = "Year", y = "Total Count", title = "Rana Muscosa Life Stage by Year", subtitle = "Data from: Knapp, R.A. 2020.") +
  scale_color_manual(values = c("skyblue1","slateblue1","royalblue4")) +
  #facet_wrap(~ amphibian_life_stage) +
  theme_minimal() +
  scale_fill_discrete(name = "Life Stage") +
  scale_fill_manual(values = c("dodgerblue1", "mediumaquamarine", "darkorchid3"))

# NOTE: I am keeping the cord flip here as to allow space to see the full lake ID without them getting compressed
# I acknolege that changing the "total count" from vertical to hortizontal is not quite ideal
graph4 <- ggplot(data = sierra_top5, aes(x= fct_reorder(lake_id, total_count), y = total_count)) + 
  geom_col(aes(fill = lake_id)) + 
  labs(x = 'Lake ID', y = 'Total Count') +
  coord_flip() + #flips so columns go from vertical alignment to horizontal alginment
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "Lake ID", y = "Total Count", title = "Top Lakes Adult/Subadult Pop.", subtitle = "Data from: Knapp, R.A. 2020.") +
  scale_fill_manual(values = c("slategrey", "steelblue4", "slateblue", "steelblue3", "slateblue4"))


plot_grid(graph3, graph4, labels = c("A.", "B."), label_size = 20)
```

**Figure 3: ** **Plot A** on the left illustrates the distribution of Rana Muscosa by life stage in in Sierra Lakes Inventory Project. Blue indicates Rana Muscosa at the tadpole lifestage, green indicating the SubAdult lifestage, and red/orange in the adult lifestage. **Plot B** on the right illustrates the lakes with top 5 counts of Rana Muscosa using the metric of combined adult and subadult populations


***

## 4. Data Citation

Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3
