---
title: "Assignment 1: Task 1"
author: "Sophia Leiker"
date: "1/16/2022"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

## 1. Introduction

This report will explore amphibian abundance data recorded by the Sierra Lakes Inventory Project. From the Environmental Data Initiative repository: “The Sierra Lakes Inventory Project (SLIP) was a research endeavor that ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. We described the physical characteristics of and surveyed aquatic communities for >8,000 lentic water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows.”

The data used comes from: [Sierra Amphibians Dataset](https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3) and metadata, including methods used for data collection can be found here [The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California](https://portal.edirepository.org/nis/metadataviewer?packageid=edi.577.2).


```{r}
# Reading in the data
sierra <- read_csv(here("data", "sierra_amphibians.csv")) %>% 
  clean_names()

```

***

# Part 1

## A. Data Wrangling 
- Selecting Mountain yellow-legged frog (Rana muscosa) --> RAMU
- Finding count each year across all water bodies by life stage
- Exclude 'EggMass' level (find the total number of adult, subadult, and tadpole yellow-legged frogs observed in the entire study region by life stage and year)

```{r filtering}
# filtering to only include certain species of interest
#using lubricate to change survery_date to an actual date 
#only selecting adult, subadult and tadpole

sierra_yellow <- sierra %>% 
  filter(amphibian_species %in% c("RAMU")) %>% 
  mutate(survey_date = mdy(survey_date)) %>% 
  mutate(survey_date = year(survey_date)) %>% 
  filter(amphibian_life_stage %in% c("SubAdult", "Tadpole", "Adult"))
```

## B. Data Analysis
- Grouping by life stage and survey date

```{r}
sierra_count <- sierra_yellow %>% 
  group_by(amphibian_life_stage, survey_date) %>% 
  summarize(total_count = n())

#head(sierra_count)

```

## C. Data visualization
- SOPHIA TRY AND RECOLOR THESE ULGY COLORS

```{r}
ggplot(data = sierra_count, aes(x=survey_date, y= total_count, fill= amphibian_life_stage)) +
  geom_col() +
  labs(x = "Year", y = "Total Count", title = "Amphibian Life Stage for Rana Muscosa in Sierra Lakes Inventory Project", subtitle = "Data from: Knapp, R.A. 2020. The Sierra Lakes Inventory Project") +
  scale_color_manual(values = c("skyblue1","slateblue1","royalblue4")) +
  facet_wrap(~ amphibian_life_stage) +
  theme_minimal()
```
***

# Part 2

## A. Data Analysis
- Total counts of combined adult and sub-adult frogs in the 5 lakes with greatest total counts
- 5 columns with lake ID

```{r}
sierra_yellow2 <- sierra_yellow %>% 
  filter(amphibian_life_stage %in% c("SubAdult", "Adult")) %>% 
  group_by(lake_id) %>% 
  summarize(total_count = n())

sierra_top5 <- sierra_yellow2 %>% 
  slice_max(total_count, n = 5) %>% #the 6th value is lower than the 5th, so there are no ties between the 5th/6th
  arrange(desc(total_count)) %>% 
  mutate(lake_id = as.character(lake_id))

# Adding "Lake" to the beginning of all the lake names
sierra_top5[1, "lake_id"] <- "Lake 50183"
sierra_top5[2, "lake_id"] <- "Lake 70583"
sierra_top5[3, "lake_id"] <- "Lake 10226"
sierra_top5[4, "lake_id"] <- "Lake 41322"
sierra_top5[5, "lake_id"] <- "Lake 50219"



```

## B. Data Visualization
- SOPHIA SHORTEN TITLE
- SOPHIA SEE HOW TO CHANGE THE COLORS
```{r}
ggplot(data = sierra_top5, aes(x= fct_reorder(lake_id, total_count), y = total_count)) + 
  geom_col(aes(fill = lake_id)) + 
  labs(x = 'Lake ID', y = 'Total Count') +
  #coord_flip() + #flips so columns go from vertical alignment to horizontal alginment
  theme_minimal() +
  labs(x = "Total Count", y = "Lake ID", title = "Combined Adult and Subadult Rana Muscosa in Lakes with Top 5 Total Observed Counts", subtitle = "Data from: Knapp, R.A. 2020. The Sierra Lakes Inventory Project")
```



***

## X. Data Citation

Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3
